Eye in a Disk: eyeIntegration human pan-eye and body transcriptome database version 1.00

### 2019-01-30 manuscript.Rmd

## RNA-seq is the predominant technology for deciphering transcriptomes
From anterior to posterior, the human eye is composed of the cornea, lens, retina, retinal pigment epithelium (RPE), and choroid. The differentiation, development, and function of these tissues is mediated through tissue and temporal specific transcript and gene expression patterns, also known as the transcriptome. Today, RNA-sequencing (RNA-seq) is the predominant technology for quantifying the transcriptome. Analysis of the transcriptome across tissue, time, and perturbation allows researchers to decipher the genetic controls of eye development and function. To this end researchers have used a wide variety of tissue types to assess gene function, including primary tissue (fetal and post-mortem), differentiated stem cells, immortalized cell lines, and most recently, organoids. These tissue types have been deeply sequenced across the cornea [@chenIdentificationNovelMolecular2013a; @chngHighThroughputGene2013a; @chungTranscriptomicProfilingPosterior2017; @fraustoTranscriptomicAnalysisCultured2016a; @kabzaCollagenSynthesisDisruption2017; @ouyangWNT7APAX6Define2014a; @songDirectedDifferentiationHuman2016a], lens [@hanWnt5aContributesDifferentiation2018], retina [@aldiriDynamicEpigeneticLandscape2017a; @farkasTranscriptomeAnalysesHuman2013b; @hoshinoMolecularAnatomyDeveloping2017; @kaewkhawTranscriptomeDynamicsDeveloping2015; @kaewkhawTreatmentParadigmsRetinal2016; @liComprehensiveAnalysisGene2014b; @mustafiTranscriptomeAnalysisReveals2016; @pinelliAtlasGeneExpression2016; @whitmoreTranscriptomicAnalysisNasal2014b], and RPE (choroid) [@darrowDeletionDXZ4Human2016a; @harenzaTranscriptomicProfiling392017; @huIdentificationMiRNASignatures2012b; @liComprehensiveAnalysisGene2014b; @nozawaSAFARegulatesInterphase2017; @obersteinCellularResponsesHuman2017; @pengEngineeringBloodRetinalBarrier2013; @radekeRestorationMesenchymalRetinal2015a; @sainiNicotinamideAmelioratesDisease2017; @samuelAppropriatelyDifferentiatedARPE192017; @santaguidaAneuploidyinducedCellularStresses2015a; @shaoSpontaneousGenerationNovel2017; @shihRestorationMesenchymalRPE2017; @smithRetinalPigmentEpithelial2017; @stevensonGiantinknockoutModelsReveal2017; @tresiniCoreSpliceosomeTarget2015; @whewaySiRNAbasedFunctionalGenomics2015; @whitmoreTranscriptomicAnalysisNasal2014b; @auCharacterizationLincRNAExpression2017].

@@

>>>>

## RNA-seq is the predominant technology for deciphering transcriptomes

From anterior to posterior, the human eye is composed of the cornea, lens, retina, retinal pigment epithelium (RPE), and choroid. The differentiation, development, and function of these tissues is mediated through tissue and temporal specific transcript and gene expression patterns, also known as the transcriptome. Today, RNA-sequencing (RNA-seq) is the predominant technology for quantifying the transcriptome. Analysis of the transcriptome across tissue, time, and perturbation allows researchers to decipher the genetic controls of eye development and function. To this end researchers have used a wide variety of tissue types to assess gene function, including primary tissue (fetal and post-mortem), differentiated stem cells, immortalized cell lines, and most recently, organoids. These tissue types have been deeply sequenced across the cornea [@chenIdentificationNovelMolecular2013a; @chngHighThroughputGene2013a; @chungTranscriptomicProfilingPosterior2017; @fraustoTranscriptomicAnalysisCultured2016a; @kabzaCollagenSynthesisDisruption2017; @ouyangWNT7APAX6Define2014a; @songDirectedDifferentiationHuman2016a], lens [@hanWnt5aContributesDifferentiation2018], retina [@aldiriDynamicEpigeneticLandscape2017a; @farkasTranscriptomeAnalysesHuman2013b; @hoshinoMolecularAnatomyDeveloping2017; @kaewkhawTranscriptomeDynamicsDeveloping2015; @kaewkhawTreatmentParadigmsRetinal2016; @liComprehensiveAnalysisGene2014b; @mustafiTranscriptomeAnalysisReveals2016; @pinelliAtlasGeneExpression2016; @whitmoreTranscriptomicAnalysisNasal2014b], and RPE (choroid) [@darrowDeletionDXZ4Human2016a; @harenzaTranscriptomicProfiling392017; @huIdentificationMiRNASignatures2012b; @liComprehensiveAnalysisGene2014b; @nozawaSAFARegulatesInterphase2017; @obersteinCellularResponsesHuman2017; @pengEngineeringBloodRetinalBarrier2013; @radekeRestorationMesenchymalRetinal2015a; @sainiNicotinamideAmelioratesDisease2017; @samuelAppropriatelyDifferentiatedARPE192017; @santaguidaAneuploidyinducedCellularStresses2015a; @shaoSpontaneousGenerationNovel2017; @shihRestorationMesenchymalRPE2017; @smithRetinalPigmentEpithelial2017; @stevensonGiantinknockoutModelsReveal2017; @tresiniCoreSpliceosomeTarget2015; @whewaySiRNAbasedFunctionalGenomics2015; @whitmoreTranscriptomicAnalysisNasal2014b; @auCharacterizationLincRNAExpression2017].

@@

# Funding

This research was supported by the Intramural Research Program of the National Eye Institute, National Institutes of Health.
